---
title: "p8105_hw3_co2616"
output: github_document
date: "2025-10-13"
---

```{r}
library(tidyverse)
```

## Problem 1

loading instacart data

```{r}
library(p8105.datasets)
data("instacart")
```

The `instacart` dataset has `r nrow(instacart)` observations and `r ncol(instacart)` variables and tells us the order and product details of online grocery orders via instacart in 2017. This dataset contains key variables such as `product_name` which shows the name of the food products ordered, `department` which describes the department the food product is categorized under, `add_to_cart_order` which tells the order in which the food product was added to the online cart, and `order_hour_of_day` which tells the hour of the day in which the order was placed. The average order number in which a food product was added to the cart is `r mean(pull(instacart, add_to_cart_order))` and the median hour of the day in which a order was placed is `r median(pull(instacart, order_hour_of_day))`.

```{r}
instacart %>% 
  group_by(order_hour_of_day) %>% 
  ggplot(aes(x = order_hour_of_day)) +
  geom_histogram()
```

**How many aisles are there, and which aisles are the most items ordered from?**

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n())
```

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n()) %>% 
  filter(min_rank(desc(n_obs)) <  4)
```

There are 134 aisles and the top 3 aisles with the most items ordered are fresh fruits (150,473 items), fresh vegetables (150,609 items), and packaged vegetables fruits (78,493 items).

**Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.**

```{r}
instacart %>% 
  group_by(aisle, department_id) %>% 
  summarize(n_obs = n()) %>% 
  filter(n_obs > 10000) %>% 
  ggplot(aes(x = aisle, y = n_obs)) +
  geom_col()
```

**Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.**

```{r}
 instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarise(n_obs = n()) %>%
  group_by(aisle) %>% 
  mutate(product_ranking = min_rank(desc(n_obs))) %>% 
  filter(product_ranking < 4) %>% 
  knitr::kable()
```

**Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).**

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarise(
    mean_hour = mean(order_hour_of_day, na.rm = TRUE)
  ) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>% 
  knitr::kable()
```

## Problem 2

loading and tidying zillow datasets

```{r}
zip_df = 
   read_csv("zillow_data/Zip Codes.csv", na = c("NA", ".", "")) %>%
  janitor::clean_names() %>% 
  mutate(
    borough = case_match(
      county,
      "Bronx" ~ "Bronx",
      "Kings" ~ "Brooklyn",
      "New York" ~ "Manhattan",
      "Queens" ~ "Queens",
      "Richmond" ~ "Staten Island")) %>% 
  select(-county_fips, -state_fips, -county_code, -county, -file_date) %>%
  relocate(zip_code)

zori_df = 
   read_csv("zillow_data/Zip_zori.csv", na = c("NA", ".", "")) %>%
  janitor::clean_names() %>% 
  rename(
    zip_code = region_name
  ) %>%
  pivot_longer(
    cols = x2015_01_31:x2024_08_31,
    names_to = "date",
    values_to = "rental_price",
    names_prefix = "x") %>% 
  select(-region_type, -state_name, -metro, -county_name, -size_rank) %>% 
  relocate(zip_code)

nyc_zillow_df = 
  left_join(zip_df, zori_df, by = "zip_code")
```

**There are 116 months between January 2015 and August 2024. How many ZIP codes are observed 116 times? How many are observed fewer than 10 times? Why are some ZIP codes are observed rarely and others observed in each month?**

```{r}
nyc_zillow_df %>%
  group_by(zip_code) %>%
  summarise(n_obs = n()) %>% 
  filter(n_obs == 116)

nyc_zillow_df %>%
  group_by(zip_code) %>%
  summarise(n_obs = n()) %>% 
  filter(n_obs < 10)
```

Between January 2015 and August 2024, there are 147 ZIP codes observed 116 times and there are 171 ZIP codes observed less than 10 times. Some ZIP codes are observed rarely and others observed in each month because the Zillow website shows available homes for sale or rent. So, if a home was sold or rented out, it was probably removed from the website along with the associated ZIP Code going forward.Thus the ZIP code for that home would longer show up in the subsequent months on the Zillow website.

**Create a reader-friendly table showing the average rental price in each borough and year (not month). Comment on trends in this table.**

```{r}
nyc_zillow_df %>% 
  drop_na(rental_price) %>% 
  separate(
    date, into = c("year", "month", "day")) %>% 
  group_by(borough, year) %>% 
  summarize(
  mean_rental_price = mean(rental_price, na.rm = TRUE)
  ) %>% 
  pivot_wider(
    names_from = borough,
    values_from = mean_rental_price
  )
```
I noticed the average rental prices under Staten Island from the years 2015 through 2019 are missing. Manhattan has the highest average rental prices compared to the other boroughs across all years and the Bronx has the lowest average rental prices. The average rental prices in Queens has stayed fairly consistent across the years compared to the other boroughs.

**Make a plot showing NYC Rental Prices within ZIP codes for all available years. Your plot should facilitate comparisons across boroughs. Comment on any significant elements of this plot.**

```{r}
nyc_zillow_df %>% 
  ggplot(aes(x = rental_price)) +
  geom_histogram() +
  facet_grid(. ~ borough) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

**Compute the average rental price within each ZIP code over each month in 2023. Make a reader-friendly plot showing the distribution of ZIP-code-level rental prices across boroughs; put differently, your plot should facilitate the comparison of the distribution of average rental prices across boroughs. Comment on this plot.**
```{r}
nyc_zillow_df %>% 
  drop_na(rental_price) %>% 
  separate(
    date, into = c("year", "month", "day")) %>% 
  filter(year == 2023) %>% 
  group_by(borough, zip_code, month) %>% 
  summarize(
  mean_rental_price = mean(rental_price, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = month, y = mean_rental_price)) +
  geom_violin() +
  facet_grid(borough ~ .) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

**Combine the two previous plots into a single graphic, and export this to a results folder in your repository**
```{r}

```

